They're Clutching up! Team Momentum in Round-Based Esports

### 2022-07-31 presentation.qmd
move round 1 win pct
formalize the momentum analysis

@@

simulate_post_streak_prob <- function(sims = 1000, ...) {
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}
# prob_n6_k2 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 2, p = 0.5)
prob_n6_k3 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 3, p = 0.5)
prob_n11_k3 <- simulate_post_streak_prob(sims = 10000, n = 8, k = 3, p = 0.5)
prob_n11_k3 <- simulate_post_streak_prob(sims = 10000, n = 11, k = 3, p = 0.5)
prob_n10_k1 <- simulate_post_streak_prob(sims = 10000, n = 10, k = 1, p = 0.5)

cod_round_streaks <- cod_rounds |> 
  # filter(round <= 6) |> 
  group_by(year, sheet, series, team) |> 
  mutate(
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA)
  ) |> 
  ungroup() |>
  select(
    year,
    sheet,
    series,
    team,
    is_offense,
    round,
    win_round,
    won_prior_round,
    won_prior_round2,
    won_prior_round3
  )

cod_round_streaks_after_b2b2b <- cod_round_streaks |> 
  filter(!is.na(won_prior_round2), !is.na(won_prior_round3)) |> 
  count(win_round, won_prior_round, won_prior_round2, won_prior_round3, sort = TRUE)

cod_round_win_prop_after_b2b2b_wins <- cod_round_streaks_after_b2b2b |> 
  filter(won_prior_round, won_prior_round2, won_prior_round3) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  filter(win_round) |> 
  select(n, total, prop)
```

```{r}
#| label: cod_round_win_prop_after_b2b2b_wins-eval
#| eval: true
prob_n6_k3 <- 0.3838163
cod_round_win_prop_after_b2b2b_wins <- tibble(
  n = 621,
  total = 1150
) |> 
  mutate(prop = n / total)
```

```{r}
#| label: summarize_and_compare_win_prop_after_wins
## https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422#app2 suggests that we should look at non-overlapping streaks.
## "To avoid this problem, we examine only the frequency of winning game three after having won games one and two and the frequency of winning game six after having won games four and five". They sort of deduced that win %s > 50% in the third game might constitute momentum, but didn't make strong statements.
summarize_cod_win_prop_after_x_wins <- function(round, ...) {
  cod_round_streaks |> 
    filter(round == !!round) |> 
    count(win_round, won_prior_round, won_prior_round2, ..., sort = TRUE) |> 
    relocate(n) |> 
    drop_na(last_col()) |> 
    filter(won_prior_round, won_prior_round2, ...) |> 
    mutate(
      total = sum(n),
      prop = n / total
    ) |> 
    filter(win_round) |> 
    transmute(round = !!round, n, total, prop)
}

summarize_cod_win_prop_after_b2b_wins <- summarize_cod_win_prop_after_x_wins
summarize_cod_win_prop_after_b2b2b_wins <- function(round) {
  summarize_cod_win_prop_after_x_wins(round = round, won_prior_round3)
}

summarize_and_compare_cod_win_prop_after_wins <- function(ri, add, f) {
  ra <- f(ri)
  rb <- f(ri + add)
  bind_rows(ra, rb)
}

summarize_and_compare_cod_win_prop_after_b2b_wins <- partial(
  summarize_and_compare_cod_win_prop_after_wins,
  add = 3,
  f = summarize_cod_win_prop_after_b2b_wins,
  ... = 
)

summarize_and_compare_cod_win_prop_after_b2b2b_wins <- partial(
  summarize_and_compare_cod_win_prop_after_wins,
  add = 4,
  f = summarize_cod_win_prop_after_b2b2b_wins,
  ... = 
)

cod_win_prop_after_b2b_wins <- seq(3, 8, by = 1) |> 
  map_dfr(
    summarize_and_compare_cod_win_prop_after_b2b_wins,
    .id = 'pair_idx'
  )

cod_win_prop_after_b2b2b_wins <- seq(4, 7, by = 1) |> 
  map_dfr(
    summarize_and_compare_cod_win_prop_after_b2b2b_wins,
    .id = 'pair_idx'
  )
```

Round win % after a streak of 3 round wins:

-   Expected: **`r 100 * round(prob_n6_k3, 3)`%**
-   Actual: **`r 100 * round(cod_round_win_prop_after_b2b2b_wins$prop, 3)`%** (`r cod_round_win_prop_after_b2b2b_wins$n` / `r cod_round_win_prop_after_b2b2b_wins$total`)

## Is planting at one site better than the other?

>>>>

::: fragment
[Miller and Sanjurjo (2018)](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2627354) outline a framework for quantifying post-streak success
:::

::: fragment
Looking at streaks of 3
:::

@@

simulate_post_streak_prob <- function(sims = 10000, seed = 42, ...) {
  withr::local_seed(seed)
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}

summarize_cod_win_prop_after_x_wins <- function(kind = c('b2b', 'b2b2b')) {
  kind <- match.arg(kind)
  is_b2b <- kind == 'b2b'
  if (is_b2b) {
    k <- 2
  } else {
    k <- 3
  }

  probs_nx_k <- tibble(
    n_rounds = 7L:11L
  ) |> 
    mutate(
      trial_p = 6 / n_rounds,
      expected_poststreak_p = map2_dbl(
        n_rounds, trial_p, 
        ~simulate_post_streak_prob(n = ..1, k = !!k, p = ..2)
      )
    )

  cod_round_win_prop_after_x_wins <- if (is_b2b) {
    cod_round_streaks_after_x <- cod_round_streaks |> 
      filter(!is.na(won_prior_round2)) |> 
      count(n_rounds, win_round, won_prior_round, won_prior_round2, sort = TRUE)

     cod_round_streaks_after_x |> 
      filter(won_prior_round, won_prior_round2) 
  } else {
    cod_round_streaks_after_x <- cod_round_streaks |> 
      filter(!is.na(won_prior_round2), !is.na(won_prior_round3)) |> 
      count(n_rounds, win_round, won_prior_round, won_prior_round2, won_prior_round3, sort = TRUE)

     cod_round_streaks_after_x |> 
      filter(won_prior_round, won_prior_round2, won_prior_round3) 
  }

  cod_round_win_prop_after_x_wins |> 
    group_by(n_rounds) |> 
    mutate(
      total = sum(n),
      prop = n / total
    ) |> 
    ungroup() |> 
    filter(win_round) |> 
    select(n_rounds, n, total, prop) |> 
    arrange(n_rounds) |> 
    inner_join(
      probs_nx_k,
      by = 'n_rounds'
    ) |> 
    nest(data = -c(n_rounds)) |> 
    mutate(
      gt_p_value = map_dbl(
        data, 
        ~binom.test(.x$n, .x$total, .x$expected_poststreak_p, alternative = 'greater') |> 
          broom::tidy() |>
          pull(p.value)
      ),
      lt_p_value = map_dbl(
        data, 
        ~binom.test(.x$n, .x$total, .x$expected_poststreak_p, alternative = 'less') |> 
          broom::tidy() |>
          pull(p.value)
      )
    ) |> 
    unnest(data)
}

cod_round_win_prop_after_b2b_wins <- summarize_cod_win_prop_after_x_wins('b2b')
cod_round_win_prop_after_b2b2b_wins <- summarize_cod_win_prop_after_x_wins('b2b2b')
```

```{r}
#| label: cod_round_win_prop_after_b2b2b_wins-eval
#| eval: true
# cod_round_win_prop_after_b2b_wins |> clipr::write_clip() |> datapasta::tribble_paste()
cod_round_win_prop_after_b2b_wins <- tibble::tribble(
  ~n_rounds,   ~n, ~total,             ~prop,          ~trial_p, ~expected_poststreak_p,       ~gt_p_value,          ~lt_p_value,
         7L, 262L,   328L, 0.798780487804878, 0.857142857142857,      0.786573045059608,   0.3220219536276,    0.725316732579212,
         8L, 265L,   421L, 0.629453681710214,              0.75,      0.648791374811152, 0.811481245561487,    0.217078977854197,
         9L, 275L,   577L, 0.476603119584055, 0.666666666666667,      0.557203226372085, 0.999956606998609, 6.11343021400038e-05,
        10L, 257L,   628L, 0.409235668789809,               0.6,      0.483529268542724, 0.999920968231117, 0.000108871179188732,
        11L, 317L,   807L, 0.392812887236679, 0.545454545454545,       0.42596043311709, 0.974133848646825,   0.0304921375685209
  )
# cod_round_win_prop_after_b2b2b_wins |> clipr::write_clip() |> datapasta::tribble_paste()
cod_round_win_prop_after_b2b2b_wins <- tibble::tribble(
  ~n_rounds,   ~n, ~total,             ~prop,          ~trial_p, ~expected_poststreak_p,       ~gt_p_value,         ~lt_p_value,
         7L, 149L,   201L, 0.741293532338308, 0.857142857142857,      0.757307821042062, 0.732554638281868,   0.323308922871945,
         8L, 126L,   203L, 0.620689655172414,              0.75,      0.618543392752532, 0.505988058731931,   0.551527754692743,
         9L,  98L,   209L,   0.4688995215311, 0.666666666666667,      0.526500674199098, 0.958718294702725,   0.055063286515911,
        10L,  74L,   208L, 0.355769230769231,               0.6,      0.446500557074851, 0.996786203815779, 0.00490308277112825,
        11L,  90L,   276L, 0.326086956521739, 0.545454545454545,      0.387843835660101, 0.985650077371605,  0.0196488641002811
  )
```

::: fragment
There is no evidence for momentum, for both streaks of 2 and 3. In fact, for series that go to 9, 10, or 11 rounds, there is evidence that the streaking teams perform significantly worse than expected in the post-streak round.
:::

```{r}
#| label: chance_summarize_and_compare_win_prop_after_wins
#| eval: false
## https://content.iospress.com/articles/journal-of-sports-analytics/jsa200422#app2 suggests that we should look at non-overlapping streaks.
## "To avoid this problem, we examine only the frequency of winning game three after having won games one and two and the frequency of winning game six after having won games four and five". They sort of deduced that win %s > 50% in the third game might constitute momentum, but didn't make strong statements.
## Personally, I don't really like this approach. It needs to account for the number of games played.
chance_summarize_cod_win_prop_after_x_wins <- function(round, ...) {
  cod_round_streaks |> 
    filter(round == !!round) |> 
    count(win_round, won_prior_round, won_prior_round2, ..., sort = TRUE) |> 
    relocate(n) |> 
    drop_na(last_col()) |> 
    filter(won_prior_round, won_prior_round2, ...) |> 
    mutate(
      total = sum(n),
      prop = n / total
    ) |> 
    filter(win_round) |> 
    transmute(round = !!round, n, total, prop)
}

chance_summarize_cod_win_prop_after_b2b_wins <- chance_summarize_cod_win_prop_after_x_wins
chance_summarize_cod_win_prop_after_b2b2b_wins <- function(round) {
  chance_summarize_cod_win_prop_after_x_wins(round = round, won_prior_round3)
}

chance_summarize_and_compare_cod_win_prop_after_wins <- function(ri, add, f) {
  ra <- f(ri)
  rb <- f(ri + add)
  bind_rows(ra, rb)
}

chance_summarize_and_compare_cod_win_prop_after_b2b_wins <- partial(
  chance_summarize_and_compare_cod_win_prop_after_wins,
  add = 3,
  f = summarize_cod_win_prop_after_b2b_wins,
  ... = 
)

chance_summarize_and_compare_cod_win_prop_after_b2b2b_wins <- partial(
  chance_summarize_and_compare_cod_win_prop_after_wins,
  add = 4,
  f = summarize_cod_win_prop_after_b2b2b_wins,
  ... = 
)

chance_cod_win_prop_after_b2b_wins <- seq(3, 8, by = 1) |> 
  map_dfr(
    chance_summarize_and_compare_cod_win_prop_after_b2b_wins,
    .id = 'pair_idx'
  )

chance_cod_win_prop_after_b2b2b_wins <- seq(4, 7, by = 1) |> 
  map_dfr(
    chance_summarize_and_compare_cod_win_prop_after_b2b2b_wins,
    .id = 'pair_idx'
  )
```

## Is planting at one site better than the other ([COD SnD]{.cod})?
